---
title: "pstat131-hw2"
author: "Sissi Shen"
date: "2022-10-15"
output: html_document
---

```{r}
abalone = read.csv(file='abalone.csv')
```

Question 1:\n
```{r}
abalone <- abalone %>%
  mutate(age = rings + 1.5)
hist(abalone$age)
```
\n
From the histogram, we can tell that the distribution of abalone's age is skewed to the left, which indicates that most abalone in our data sample are relatively young.
\n

Question 2:
```{r}
set.seed(1029)
abalone_split <- initial_split(abalone, prop=0.8, strata = age)
abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)
```
\n

Question 3:
```{r}
abalone <- abalone %>% select(-rings)
```
\n
We should not include the variable "rings" because it is bascially the outcome "age" that we want to predict. 
\n
```{r}
simple_abalone_recipe <- recipe(age ~ ., data = abalone_train)
simple_abalone_recipe
```
```{r}
abalone_recipe <- recipe(age ~ ., data = abalone_train) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_interact(terms = ~ starts_with("type"):shucked_weight) %>%
  step_interact(terms = ~ longest_shell:diameter) %>%
  step_interact(terms = ~ shucked_weight:shell_weight) %>%
  step_center(all_numeric_predictors()) %>%
  step_scale(all_numeric_predictors())
```
\n

Question 4:
```{r}
lm_model <- linear_reg() %>%
  set_engine("lm")
```
\n

Question 5:
```{r}
lm_wflow <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(abalone_recipe)
```
\n

Question 6:
```{r}
lm_fit <- fit(lm_wflow, abalone_train)
lm_fit %>%
  extract_fit_parsnip()%>%
  tidy()
```
```{r}
H_abalone <- data.frame(type = "F", longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1)
predict(lm_fit, H_abalone)
```
\n
From the model's prediction, the age of our hypothetical female abalone should be 25.87 years old.
\n

Question 7:
```{r}
abalone_train_res <- predict(lm_fit, new_data = abalone_train %>% select(-age))
abalone_train_res <- bind_cols(abalone_train_res, abalone_train %>% select(age))
abalone_train_res %>%
  head()
rmse(abalone_train_res, truth = age, estimate = .pred)
abalone_metrics <- metric_set(rmse, rsq, mae)
abalone_metrics(abalone_train_res, truth = age, estimate = .pred)
```
\n
The model has a root mean squared error (RMSE) of 2.138, an $R^2$ of 0.5575 and a mean absolute error (MAE) of 1.542. An $R^2$ value of 0.5575 means that the model can explain 55.75% of the variance of the observed data. 